---
title: "Advanced Data Visualizations"
author: "Jeff Goseland"
date: last-modified
subtitle: "West - Week 3"
format: 
  html:
    df-print: paged
    embed-resources: true
---

```{css echo=FALSE}
code.r{ /* Code block */
     background-color: lightyellow;
     border: 1px solid black;
} 

pre{ /* Output Block */
     background-color: #F1F1F1;
     border: 1px solid black;
} 

```

## Week 3 Homework

## Part A

```{r warning=FALSE, message=FALSE}
#loads needed libraries
suppressMessages({
  library(tidyverse)
  library(RColorBrewer)
  library(sf)
  library(ggmap)
})
```

#### 1. Load spacial data from a shapefile into R

```{r}
filename <- '../data/School_Boundaries.shp'

school_boundaries_shp <- st_read(filename, stringsAsFactors = FALSE)  
```

```{r}
summary(school_boundaries_shp)
```

```{r}
school_boundaries_shp[1,0]
```

```{r}
plot(school_boundaries_shp)
```

```{r}
plot(school_boundaries_shp$geometry)

```

#### 2. Create a spatial data frame from a table of latitude and longitudes

-   Include the code you used to load the csv;
-   Include the code you used to convert it to a spatial data frame;
-   Include the output of the summary() command run on the spatial data frame; and
-   Create a static map with the colors of the features reflecting some variable.

```{r}
school_boundaries_points <- read.csv("../data/Parks_Locations_and_Features.csv")
```
